---
title: "Portfolio 1: Pixar Movie Ratings"
---

### Goal 

### Data

### Product

### Interpretation


```{r load-packages, warning = FALSE, message=FALSE}
library(tidyverse) 
library(tidyr)
library(tidytuesdayR)
library(pixarfilms)
```


```{r read-data}
tuesdata <- tidytuesdayR::tt_load('2025-03-11')
pixar_films <- tuesdata$pixar_films
public_response <- tuesdata$public_response
box_office <- pixarfilms::box_office
```

## According to each rating system, which Pixar movie is the best?

```{r RT-fav}
public_response %>%
  count(film, rotten_tomatoes) %>%
  arrange(desc(rotten_tomatoes))
```

```{r MC-fav}
public_response %>%
  count(film, metacritic) %>%
  arrange(desc(metacritic))
```


```{r CC-fav}
public_response %>%
  count(film, critics_choice) %>%
  arrange(desc(critics_choice))
```

## Do audiences and critics tend to agree?

```{r scatter-RT-CC, warning=FALSE}
ggplot(
  data = public_response,
  mapping = aes(
    x = rotten_tomatoes,
    y = critics_choice
  )
) +
  geom_point() + 
  
  labs(x = "Rotten Tomatoes Score",
       y = "Critic's Choice") 
```

```{r RT-CC-cor}
cor.test(public_response$rotten_tomatoes, public_response$critics_choice)
```

## Which Pixar movie had the highest budget?

```{r budget}
box_office %>%
  count(film, budget) %>%
  arrange(desc(budget))
```

## Which Pixar movies performed the best at the box office?

```{r box-US-CA}
box_office %>%
  count(film, box_office_us_canada) %>%
  arrange(desc(box_office_us_canada))
```

```{r box-other}
box_office %>%
  count(film, box_office_other) %>%
  arrange(desc(box_office_other))
```

```{r box-global}
box_office %>%
  count(film, box_office_worldwide) %>%
  arrange(desc(box_office_worldwide))
```

## Do higher-budget movies receive higher audience ratings? Critic ratings?

```{r join-boxoffice-ratings}
box_office_ratings <- full_join(box_office, public_response,
  by = "film"
)
box_office_ratings
```


```{r scatter-budget-RT, warning=FALSE}
ggplot(
  data = box_office_ratings,
  mapping = aes(
    x = budget,
    y = rotten_tomatoes
  )
) +
  geom_point() + 
  
  labs(x = "Budget",
       y = "Rotten Tomatoes Score") 
```

```{r cor-budget-RT}
cor.test(box_office_ratings$budget, box_office_ratings$rotten_tomatoes)
```

```{r scatter-budget-CC, warning=FALSE}
ggplot(
  data = box_office_ratings,
  mapping = aes(
    x = budget,
    y = critics_choice
  )
) +
  geom_point() + 
  
  labs(x = "Budget",
       y = "Critic's Choice") 
```

```{r cor-budget-CC}
cor.test(box_office_ratings$budget, box_office_ratings$critics_choice)
```

## Do box office numbers correspond with audience ratings? Critics' ratings?

```{r scatter-BO-RT, warning=FALSE}
ggplot(
  data = box_office_ratings,
  mapping = aes(
    x = box_office_worldwide,
    y = rotten_tomatoes
  )
) +
  geom_point() + 
  
  labs(x = "Worldwide box office",
       y = "Rotten Tomatoes") 
```

```{r scatter-BO-CC, warning=FALSE}
ggplot(
  data = box_office_ratings,
  mapping = aes(
    x = box_office_worldwide,
    y = critics_choice
  )
) +
  geom_point() + 
  
  labs(x = "Worldwide box office",
       y = "Critics' Choice") 
```

