---
title: "Portfolio 3: Changes In Perceptions of Teens Over Time, cont."
---

## Overview:

#### Goal:

Using the same data from my previous portfolio piece, I want to condense the information about changes in views of teens over time into one figure for my conference poster. Because there was not a significant effect of condition, I will collapse across conditions and display just the overall main effect of time.

#### Data:

Using the same dataset from portfolio piece 2, for this project I will focus on the pre- and post-test scores on stereotypes about teens' antisocial behavior, normalized externalizing behavior, risk-taking behavior, academic engagement, emotional responsibility, positive approach to life, positive family influence, negative autonomy, and peer influence. I did a (somewhat inefficient) wide-to-long transformation in this project to combine the pre- and post-test scores into single variables to be grouped by a separate stereotype variable. This transformation allowed me to combine changes in views of different stereotypes into a single plot.

#### Product:

My initial attempt at a main effects plot, followed by a cleaner, prettier final version for the poster.

#### Interpretation:

On average, participants' views of teens improved over the course of the study, with significant effects of time on all but two stereotypes.

##### Sources:

<https://stackoverflow.com/questions/62385588/main-effect-plot-with-ggplot2>

<https://stackoverflow.com/questions/17180115/manually-setting-group-colors-for-ggplot2>

------------------------------------------------------------------------

### Loading packages

```{r load-packages, message=FALSE, warning=FALSE}
library(haven)
library(tidyverse)
library(car)
library(rstatix)
```

### Reading in data

```{r read-data}
counterstereotype_data <- read_sav("/Users/sophieboyd/Desktop/Research/Buchanan/SRA/CollegeData.sav")
```

### Subsetting the data to work only with the necessary variables + filtering out participants with missing information on ExperimentalTreatment (condition)

```{r filter-data}
counterstereotype_filtered <- counterstereotype_data %>%
  select(RespID, Gender, Ethnicity, Race, Age, ExperimentalTreatment,  VoT_MS_Antisocial,VOT_P2EB_ANTISOCIAL, VoT_MS_NEB, VOT_P2EB_NEB, VoT_MS_RT, VOT_P2EB_RT, VoT_MS_AE, VOT_P2AE, VoT_MS_ER, VOT_P2PA_ER, VoT_MS_PAL, VOT_P2PA_PATL, VoT_MS_PFI, VOT_P2_PFI, VoT_MS_NA, VOT_P2_NA, VoT_MS_PI, VOT_P2_PI) %>%
  filter(!is.na(ExperimentalTreatment))
```

### Wide-to-long transformation to get one pre-test variable and one post-test variable

```{r wide-to-long}
counterstereotype_long <- counterstereotype_filtered %>%
  pivot_longer(
    cols = starts_with(c("VoT_MS", "VOT_P2")),
    names_to = "time_stereotype",
    values_to = "viewsofteens")
```

### Creating time (pre vs. post) grouping variable

```{r time-var}
counterstereotype_long <- counterstereotype_long %>%
  mutate(time = ifelse(startsWith(time_stereotype, "VoT_MS"), 1, 2))

counterstereotype_long$timelabel <- factor(counterstereotype_long$time,
                                                      labels = c("Pre-Test", "Post-Test"))
```

### Creating stereotype grouping variable

```{r stereotype-var}
counterstereotype_long <- counterstereotype_long %>%
  mutate(stereotype = ifelse(time_stereotype == "VoT_MS_Antisocial", "antisocial",
                             ifelse(time_stereotype == "VOT_P2EB_ANTISOCIAL", "antisocial",
                                    ifelse(time_stereotype == "VoT_MS_NEB", "norm_ext",
                                           ifelse(time_stereotype == "VOT_P2EB_NEB", "norm_ext",
                                                  ifelse(time_stereotype == "VoT_MS_RT", "risk",
                                                         ifelse(time_stereotype == "VOT_P2EB_RT", "risk",
                                                                ifelse(time_stereotype == "VoT_MS_AE", "academic",
                                                                       ifelse(time_stereotype == "VOT_P2AE", "academic",
                                                                              ifelse(time_stereotype == "VoT_MS_ER", "emo_resp",
                                                                                     ifelse(time_stereotype == "VOT_P2PA_ER", "emo_resp",
                                                                                            ifelse(time_stereotype == "VoT_MS_PAL", "pos_app",
                                                                                                   ifelse(time_stereotype == "VOT_P2PA_PATL", "pos_app",
                                                                                                          ifelse(time_stereotype == "VoT_MS_PFI", "pos_fam",
                                                                                                                 ifelse(time_stereotype == "VOT_P2_PFI", "pos_fam",
                                                                                                                        ifelse(time_stereotype == "VoT_MS_NA", "neg_auto",
                                                                                                                               ifelse(time_stereotype == "VOT_P2_NA", "neg_auto",
                                                                                                                                      ifelse(time_stereotype == "VoT_MS_PI", "peers",
                                                                                                                                             ifelse(time_stereotype == "VOT_P2_PI", "peers",
                                                                                                                                                    "x")))))))))))))))))))
```

### First attempt at plotting change over time

```{r time-main-effect-plot-1, warning=FALSE}
counterstereotype_long %>%
  ggplot(aes(x = timelabel, y = viewsofteens, group = stereotype, color = stereotype)) +
  stat_summary(fun=mean, geom="line") +
  stat_summary(fun=mean, geom="point") +
  theme_minimal() 
```

### Creating labels for stereotypes, including significance indicators (\*)

```{r stereotype-labels}
counterstereotype_long <- counterstereotype_long %>%
  mutate(stereotypelabel = ifelse(stereotype == "antisocial", "Antisocial Behavior*",
                                  ifelse(stereotype == "norm_ext", "Normalized Externalizing Behavior*",
                                         ifelse(stereotype == "risk", "Risk-Taking*",
                                                ifelse(stereotype == "academic", "Academic Engagement",
                                                       ifelse(stereotype == "emo_resp", "Emotional Responsibility*",
                                                              ifelse(stereotype == "pos_app", "Positive Approach to Life",
                                                                     ifelse(stereotype == "pos_fam", "Positive Family Influence*",
                                                                            ifelse(stereotype == "neg_auto", "Negative Autonomy*",
                                                                                   ifelse(stereotype == "peers", "Peer Influence*", "x"))))))))))

```

### Assigning colors to stereotype groups (pink for negative, blue for positive)

```{r colors}
counterstereotype_colors <- c("Antisocial Behavior*" = "pink2", "Normalized Externalizing Behavior*" = "deeppink", "Risk-Taking*" = "violetred", "Negative Autonomy*" = "hotpink", "Peer Influence*" = "deeppink4", "Academic Engagement" = "darkturquoise", "Emotional Responsibility*" = "deepskyblue", "Positive Approach to Life" = "dodgerblue3", "Positive Family Influence*" = "blue")
```

### Customizing display order of stereotype groups (grouped by positive vs. negative and ordered to match plot)

```{r stereotype-order}
counterstereotype_long <- counterstereotype_long %>%
  mutate(stereotype_ordered = fct_relevel(stereotypelabel, 'Peer Influence*', 'Risk-Taking*', 'Normalized Externalizing Behavior*', 'Negative Autonomy*', 'Antisocial Behavior*', 'Positive Family Influence*', 'Positive Approach to Life', 'Emotional Responsibility*', 'Academic Engagement'))
```

### Improving the plot

```{r time-main-effect-plot2, warning=FALSE}
counterstereotype_long %>%
  ggplot(aes(x = timelabel, y = viewsofteens, group = stereotype_ordered, color = stereotype_ordered)) +
  stat_summary(fun=mean, geom="line") +
  stat_summary(fun=mean, geom="point") +
  labs(x = "Time",
       y = "Views of Teens",
       title = "Average Change in Perceptions of Teens Over Time",
       color = "Attribute\n* = sig. main effect of time") +
  theme_minimal() + 
  theme(plot.title = element_text(hjust = 0.5)) +
  scale_color_manual(values=counterstereotype_colors)
```

#### Changes from first attempt:

-   Added labels
-   Customized colors to group by positive vs. negative while still differentiating between stereotypes/attributes
-   Customized order of stereotypes to group by positive/negative and to match order of plot
-   Centered title
