---
title: "Portfolio 4: Using Changes in Perceptions of Teens Over Time to Predict Behavioral Intentions"
---

## Overview:

#### Goal:

As a second part of the teen counterstereotyping project, I examined whether changes in the participants' perceptions of teens predicted their own behavioral intentions. I found that changes in perceptions of teens corresponded to positive behavioral intentions. For example, decreased perceptions of teens' antisocial behavior predicted lower intentions to engage in antisocial behavior, and increased perceptions of teens' positive family relationships predicted greater intentions to express gratitude toward a family member. The goal of this piece is to find an effective way to visualize the associations between changes in views of teens and behavioral intentions. 

#### Data:

For this piece, I will be using the same dataset from portfolio pieces 2 and 3, but with new central variables. I will be using change scores to represent changes in the participants' ratings of teens on each attribute from pre- to post-test. I will also be using measures of their behavioral intentions recorded at post-test. 

#### Product:

A series of scatterplots to display the associations between change scores and personal behavioral intentions, culminating in a single figure that combines all plots.

#### Interpretation:

Improvements in views of teens predicted positive behavioral intentions. (Decreases on negative attributes predicted lower negative behavioral intentions, and increases on positive attributes predicted higher positive intentions).

##### Sources:

https://r-graph-gallery.com/package/patchwork.html

---

### Loading packages

```{r load-packages, message=FALSE, warning=FALSE}
library(haven)
library(tidyverse)
library(janitor)
library(patchwork)
```

### Reading in data

```{r read-data}
counterstereotype_data <- read_sav("/Users/sophieboyd/Desktop/Research/Buchanan/SRA/CollegeData.sav")
```

### Subsetting the data to work only with the necessary variables + filtering out participants with missing information on ExperimentalTreatment (condition) to be consistent

```{r filter-data}
counterstereotype_filtered <- counterstereotype_data %>%
  select(RespID, Gender, Ethnicity, Race, Age, ExperimentalTreatment, Change_antisocial, Change_NEB, Change_RT, Change_AE, Change_ER, Change_PAL, Change_PFI, Change_NA, Change_PI, `@2BInt_antisocial`, `@2BInt_NEB`, `@2BInt_AE`, `@2BInt_NA`, `@2BInt_PFI`, `@2BInt_NA`,`@2BInt_PR`) %>%
  filter(!is.na(ExperimentalTreatment))
```

### Scatterplot examining antisocial behavior

```{r scatter-antisocial, warning=FALSE}
counterstereotype_filtered %>%
  ggplot(aes(x = Change_antisocial, y = `@2BInt_antisocial`)) +
  geom_jitter()
```

### Adding a line and improving the appearance

```{r scatter-antisocial-clean, warning=FALSE, message=FALSE}
counterstereotype_filtered %>%
  ggplot(aes(x = Change_antisocial, y = `@2BInt_antisocial`)) +
  geom_jitter() +
  geom_smooth(method = "lm", se = FALSE, color = "pink2") + 
  labs(x = "Change in Perception of Teens' Antisocial Behavior",
       y = 'Intentions to Engage in Antisocial Behavior' ) +
  theme_minimal()
```

### Scatterplot examining normalized externalizing behavior

```{r scatter-NEB, warning=FALSE}
counterstereotype_filtered %>%
  ggplot(aes(x = Change_NEB, y = `@2BInt_NEB`)) +
  geom_jitter()
```

### Adding a line and improving the appearance

```{r scatter-NEB-clean, warning=FALSE, message=FALSE}
counterstereotype_filtered %>%
  ggplot(aes(x = Change_NEB, y = `@2BInt_NEB`)) +
  geom_jitter() +
  geom_smooth(method = "lm", se = FALSE, color = "deeppink") + 
  labs(x = "Change in Perception of Teens' Normalized Externalizing Behavior",
       y = 'Intentions to Engage in Externalizing Behavior' ) +
  theme_minimal()
```

### Scatterplot examining positive family influence

```{r scatter-PFI, warning=FALSE}
counterstereotype_filtered %>%
  ggplot(aes(x = Change_PFI, y = `@2BInt_PFI`)) +
  geom_jitter()
```

### Adding a line and improving the appearance

```{r scatter-PFI-clean, warning=FALSE, message=FALSE}
counterstereotype_filtered %>%
  ggplot(aes(x = Change_PFI, y = `@2BInt_PFI`)) +
  geom_jitter() +
  geom_smooth(method = "lm", se = FALSE, color = "blue") + 
  labs(x = "Change in Perception of Teens' Positive Family Influence",
       y = 'Intentions to Engage Positively with Family' ) +
  theme_minimal()
```

### Scatterplot examining negative autonomy

```{r scatter-NA, warning=FALSE}
counterstereotype_filtered %>%
  ggplot(aes(x = Change_NA, y = `@2BInt_NA`)) +
  geom_jitter()
```

### Adding a line and improving the appearance

```{r scatter-NA-clean, warning=FALSE, message=FALSE}
counterstereotype_filtered %>%
  ggplot(aes(x = Change_NA, y = `@2BInt_NA`)) +
  geom_jitter() +
  geom_smooth(method = "lm", se = FALSE, color = "hotpink") + 
  labs(x = "Change in Perception of Teens' Negative Autonomy",
       y = 'Intentions to Engage in Negative Autonomy' ) +
  theme_minimal()
```

### Scatterplot examining academic engagement

```{r scatter-AE, warning=FALSE}
counterstereotype_filtered %>%
  ggplot(aes(x = Change_AE, y = `@2BInt_AE`)) +
  geom_jitter()
```

### Adding a line and improving the appearance

```{r scatter-AE-clean, warning=FALSE, message=FALSE}
counterstereotype_filtered %>%
  ggplot(aes(x = Change_AE, y = `@2BInt_AE`)) +
  geom_jitter() +
  geom_smooth(method = "lm", se = FALSE, color = "darkturquoise") + 
  labs(x = "Change in Perception of Teens' Academic Engagement",
       y = 'Intentions to Be Academically Engaged' ) +
  theme_minimal()
```

### Scatterplot examining emotional responsibility

```{r scatter-ER, warning=FALSE}
counterstereotype_filtered %>%
  ggplot(aes(x = Change_ER, y = `@2BInt_PR`)) +
  geom_jitter()
```

### Adding a line and improving the appearance

```{r scatter-ER-clean, warning=FALSE, message=FALSE}
counterstereotype_filtered %>%
  ggplot(aes(x = Change_ER, y = `@2BInt_PR`)) +
  geom_jitter() +
  geom_smooth(method = "lm", se = FALSE, color = "deepskyblue") + 
  labs(x = "Change in Perception of Teens' Emotional Responsibility",
       y = 'Intentions to Be Emotionally Responsible' ) +
  theme_minimal()
```

### Attempting to put it all together (using the patchwork package)

```{r define-plots}
p1 <- counterstereotype_filtered %>%
  ggplot(aes(x = Change_antisocial, y = `@2BInt_antisocial`)) +
  geom_jitter() +
  geom_smooth(method = "lm", se = FALSE, color = "pink2") + 
  labs(x = "Change in Perception of Teens' Antisocial Behavior*",
       y = 'Intentions to Engage in Antisocial Behavior' ) +
  theme_minimal()

p2 <- counterstereotype_filtered %>%
  ggplot(aes(x = Change_NEB, y = `@2BInt_NEB`)) +
  geom_jitter() +
  geom_smooth(method = "lm", se = FALSE, color = "deeppink") + 
  labs(x = "Change in Perception of Teens' Normalized Externalizing Behavior*",
       y = 'Intentions to Engage in Externalizing Behavior' ) +
  theme_minimal()

p3 <- counterstereotype_filtered %>%
  ggplot(aes(x = Change_NA, y = `@2BInt_NA`)) +
  geom_jitter() +
  geom_smooth(method = "lm", se = FALSE, color = "hotpink") + 
  labs(x = "Change in Perception of Teens' Negative Autonomy*",
       y = 'Intentions to Engage in Negative Autonomy' ) +
  theme_minimal()

p4 <- counterstereotype_filtered %>%
  ggplot(aes(x = Change_PFI, y = `@2BInt_PFI`)) +
  geom_jitter() +
  geom_smooth(method = "lm", se = FALSE, color = "blue") + 
  labs(x = "Change in Perception of Teens' Positive Family Influence*",
       y = 'Intentions to Engage Positively with Family' ) +
  theme_minimal()

p5 <- counterstereotype_filtered %>%
  ggplot(aes(x = Change_AE, y = `@2BInt_AE`)) +
  geom_jitter() +
  geom_smooth(method = "lm", se = FALSE, color = "darkturquoise") + 
  labs(x = "Change in Perception of Teens' Academic Engagement",
       y = 'Intentions to Be Academically Engaged' ) +
  theme_minimal()

p6 <- counterstereotype_filtered %>%
  ggplot(aes(x = Change_ER, y = `@2BInt_PR`)) +
  geom_jitter() +
  geom_smooth(method = "lm", se = FALSE, color = "deepskyblue") + 
  labs(x = "Change in Perception of Teens' Emotional Responsibility*",
       y = 'Intentions to Be Emotionally Responsible' ) +
  theme_minimal()
```

```{r combine-plots, warning=FALSE, message=FALSE, fig.width = 15, fig.height = 10}
(p1 + p2 + p3) / (p4 + p5 + p6) + plot_annotation(title = 'Improvements in Views of Teens Predict More Positive Behavioral Intentions', theme = theme(plot.title = element_text(hjust = 0.5)))
```


```{r save-final-plot, warning=FALSE, message=FALSE}
combined_plot <- (p1 + p2 + p3) / (p4 + p5 + p6) + plot_annotation(title = 'Improvements in Views of Teens Predict More Positive Behavioral Intentions', theme = theme(plot.title = element_text(hjust = 0.5)))

# ggsave("SRA_BehInt-plot.png", plot = combined_plot, width = 15, height = 10)
```

## Update: After getting feedback from my lab, I have decided to recode the change scores for negative attributes only, such that positive change scores are all indicative of improvments in views of teens.


```{r recode-neg-attributes}
counterstereotype_filtered$Change_antisocial_R <- counterstereotype_filtered$Change_antisocial * -1
counterstereotype_filtered$Change_NEB_R <- counterstereotype_filtered$Change_NEB * -1
counterstereotype_filtered$Change_NA_R <- counterstereotype_filtered$Change_NA * -1
```

```{r scatter-antisocial-2, warning=FALSE, message=FALSE}
counterstereotype_filtered %>%
  ggplot(aes(x = Change_antisocial_R, y = `@2BInt_antisocial`)) +
  geom_jitter() +
  geom_smooth(method = "lm", se = FALSE, color = "pink2") + 
  labs(x = "Change in Perception of Teens' Antisocial Behavior",
       y = 'Intentions to Engage in Antisocial Behavior' ) +
  theme_minimal()
```

```{r scatter-NEB-2, warning=FALSE, message=FALSE}
counterstereotype_filtered %>%
  ggplot(aes(x = Change_NEB_R, y = `@2BInt_NEB`)) +
  geom_jitter() +
  geom_smooth(method = "lm", se = FALSE, color = "deeppink") + 
  labs(x = "Change in Perception of Teens' Normalized Externalizing Behavior",
       y = 'Intentions to Engage in Externalizing Behavior' ) +
  theme_minimal()
```

```{r scatter-NA-2, warning=FALSE, message=FALSE}
counterstereotype_filtered %>%
  ggplot(aes(x = Change_NA_R, y = `@2BInt_NA`)) +
  geom_jitter() +
  geom_smooth(method = "lm", se = FALSE, color = "hotpink") + 
  labs(x = "Change in Perception of Teens' Negative Autonomy",
       y = 'Intentions to Engage in Negative Autonomy' ) +
  theme_minimal()
```

```{r define-plots2}
p1 <- counterstereotype_filtered %>%
  ggplot(aes(x = Change_NEB_R, y = `@2BInt_NEB`)) +
  geom_jitter() +
  geom_smooth(method = "lm", se = FALSE, color = "deeppink") + 
  labs(x = "Improvement in Perception of Teens' Normalized Externalizing Behavior",
       y = 'Intentions to Engage in Externalizing Behavior' ) +
  theme_minimal()

p2 <- counterstereotype_filtered %>%
  ggplot(aes(x = Change_NA_R, y = `@2BInt_NA`)) +
  geom_jitter() +
  geom_smooth(method = "lm", se = FALSE, color = "hotpink") + 
  labs(x = "Improvement in Perception of Teens' Negative Autonomy",
       y = 'Intentions to Engage in Negative Autonomy' ) +
  theme_minimal()

p3 <- counterstereotype_filtered %>%
  ggplot(aes(x = Change_antisocial_R, y = `@2BInt_antisocial`)) +
  geom_jitter() +
  geom_smooth(method = "lm", se = FALSE, color = "pink2") + 
  labs(x = "Improvement in Perception of Teens' Antisocial Behavior",
       y = 'Intentions to Engage in Antisocial Behavior' ) +
  theme_minimal()

p4 <- counterstereotype_filtered %>%
  ggplot(aes(x = Change_PFI, y = `@2BInt_PFI`)) +
  geom_jitter() +
  geom_smooth(method = "lm", se = FALSE, color = "blue") + 
  labs(x = "Improvement in Perception of Teens' Positive Family Influence",
       y = 'Intentions to Engage Positively with Family' ) +
  theme_minimal()

p5 <- counterstereotype_filtered %>%
  ggplot(aes(x = Change_ER, y = `@2BInt_PR`)) +
  geom_jitter() +
  geom_smooth(method = "lm", se = FALSE, color = "deepskyblue") + 
  labs(x = "Improvement in Perception of Teens' Emotional Responsibility",
       y = 'Intentions to Engage Positively with Peers' ) +
  theme_minimal()

p6 <- counterstereotype_filtered %>%
  ggplot(aes(x = Change_AE, y = `@2BInt_AE`)) +
  geom_jitter() +
  geom_smooth(method = "lm", se = FALSE, color = "darkturquoise") + 
  labs(x = "Improvement in Perception of Teens' Academic Engagement",
       y = 'Intentions to Be Academically Engaged' ) +
  theme_minimal()

```

```{r combine-plots2, warning=FALSE, message=FALSE, fig.width = 15, fig.height = 10}
(p1 + p2 + p3) / (p4 + p5 + p6) 
```

##### Trying a vertical layout to fit my poster better

```{r combine-plots3, warning=FALSE, message=FALSE, fig.width = 15, fig.height = 20}
(p1 + p4) / (p2 + p5) / (p3 + p6) 
```

```{r save-vertical-plot, warning=FALSE, message=FALSE}
combined_plot <- (p1 + p4) / (p2 + p5) / (p3 + p6) + plot_annotation(title = "Improvements in Views of Teens Predict Positive Behavioral Intentions", theme = theme(plot.title = element_text(hjust = 0.5))
)

combined_plot

# ggsave("SRA_BehInt-plot-vertical.png", plot = combined_plot, width = 15, height = 25)
```

#### Breaking up the plots to fit the poster

```{r save-split-plots, warning=FALSE, message=FALSE}
combined_plot1 <- (p1 + p4) 

combined_plot1

# ggsave("SRA_BehInt-plot1.png", plot = combined_plot1, width = 15, height = 7)

combined_plot2 <- (p2 + p5) 

# ggsave("SRA_BehInt-plot2.png", plot = combined_plot2, width = 15, height = 7)

combined_plot3 <- (p3 + p6)

# ggsave("SRA_BehInt-plot3.png", plot = combined_plot3, width = 15, height = 7)

combined_plot4 <- (p2 + p5) / (p3 + p6)

# ggsave("SRA_BehInt-plot4.png", plot = combined_plot4, width = 15, height = 15)


```

