---
title: "Portfolio 6: Simulating Parenting and Civic Engagement Data"
---

## Overview:

#### Goal:

I want to use my portfolio to share some of the analytic work I did for my Master's thesis. However, I need to keep my data private. My goal for this piece is to simulate a dataset that resembles my thesis data. I plan to use this simulated dataset in future portfolio pieces.

#### Data:

My original dataset contained youth-reported measures of parenting practices, civic engagement, and demographic information collected over three waves of a longitudinal study (labeled here as Times 1-3). Participants made up 11 different cultural/geographic groups, which were represented by effect-coded predictors in my final regression models. There was quite a bit of missing data across waves, so to simplify my simulation, I will use the most restricted subsample I had (668 participants).

Specific core variables include:

- Joint decision making at Time 1 (JDM; a measure of moderate support for autonomy)
- Independent decision making at Time 1 (IDM; a measure of extreme support for autonomy)
- Maternal warmth at Time 1 (MatWarmth)
- Paternal warmth at Time 1 (PatWarmth)
- Civic engagement at Time 2 (CE_2)
- Civic engagement at Time 3 (CE_3)

Recruitment groups include:

- Naples, Italy
- Rome, Italy
- Kenya
- Philippines
- Thailand
- Sweden
- US, African American
- US, European American
- US, Hispanic
- Colombia
- Jordan

#### Product:

A simulated dataset containing all variables necessary for the analyses I completed for my thesis. (I obtained all parameter values by running descriptive and correlational analyses with the original dataset in SPSS.)

#### Interpretation:

**Fill in after completing

----

### Load packages and data

```{r load-packages, message=FALSE, warning=FALSE}
library(tidyverse) 
library(ggplot2)
library(MASS)
```

### Creating the dataframe and simulating demographic information 

#### Age at Time 1 

```{r age}
set.seed(123)
df_civicengagement <- data.frame(
  id = 1:668,
  age = rnorm(668, mean = 15.61, sd = .898)
)
```

#### Geographic/cultural recruitment group (grouping variable + effect-coded variables)

```{r IDN}
set.seed(123)
df_civicengagement$IDN <- rep(c("Italy-Naples", "Italy-Rome", "Kenya", "Philippines", "Thailand", "Sweden", "US-AA", "US-EA", "US-H", "Colombia", "Jordan"),
  times = c(76, 84, 41, 67, 50, 33, 49, 66, 39, 74, 89),
  length.out = 668
)
```

```{r IDN-effect-coded}
df_civicengagement <- df_civicengagement %>%
  mutate(Naples = case_when(
    IDN == "Italy-Naples" ~ 1,
    IDN == "US-EA" ~ -1,
    TRUE ~ 0
  ))

df_civicengagement <- df_civicengagement %>%
  mutate(Rome = case_when(
    IDN == "Italy-Rome" ~ 1,
    IDN == "US-EA" ~ -1,
    TRUE ~ 0
  ))

df_civicengagement <- df_civicengagement %>%
  mutate(Kenya = case_when(
    IDN == "Kenya" ~ 1,
    IDN == "US-EA" ~ -1,
    TRUE ~ 0
  ))

df_civicengagement <- df_civicengagement %>%
  mutate(Philippines = case_when(
    IDN == "Philippines" ~ 1,
    IDN == "US-EA" ~ -1,
    TRUE ~ 0
  ))

df_civicengagement <- df_civicengagement %>%
  mutate(Thailand = case_when(
    IDN == "Thailand" ~ 1,
    IDN == "US-EA" ~ -1,
    TRUE ~ 0
  ))

df_civicengagement <- df_civicengagement %>%
  mutate(Sweden = case_when(
    IDN == "Sweden" ~ 1,
    IDN == "US-EA" ~ -1,
    TRUE ~ 0
  ))

df_civicengagement <- df_civicengagement %>%
  mutate(US_AA = case_when(
    IDN == "US-AA" ~ 1,
    IDN == "US-EA" ~ -1,
    TRUE ~ 0
  ))

df_civicengagement <- df_civicengagement %>%
  mutate(US_H = case_when(
    IDN == "US-H" ~ 1,
    IDN == "US-EA" ~ -1,
    TRUE ~ 0
  ))

df_civicengagement <- df_civicengagement %>%
  mutate(Colombia = case_when(
    IDN == "Colombia" ~ 1,
    IDN == "US-EA" ~ -1,
    TRUE ~ 0
  ))

df_civicengagement <- df_civicengagement %>%
  mutate(Jordan = case_when(
    IDN == "Jordan" ~ 1,
    IDN == "US-EA" ~ -1,
    TRUE ~ 0
  ))
```

#### Gender 

```{r gender}
df_civicengagement$male <- rep(c("not male", "male"),
  times = c(354, 314),
  length.out = 668
)
```

#### Maximum level of parental education (in years)

```{r max-parental-ed}
df_civicengagement$max_parental_ed <- rnorm(668, mean = 13.91, sd = 4.21)
```


### Simulating the central variables

```{r parenting-CE-sim}
set.seed(123)

n_participants <- 668

var_names <- c("JDM", "IDM", "MatWarmth", "PatWarmth", "CE_2", "CE_3")

mean_vars <- c(.96, .26, 3.46, 3.26, 2.04, 1.81)
sd_vars <- c(.40, .21, .56, .74, .86, .80)

cor_matrix_parenting_CE <- matrix(
  c(
1.0000, -.802, .129, .172, .086, .103, 
-.802, 1.0000, -.148, -.197, -.129, -.102,
.129, -.148, 1.0000, .551, .056, -.002,
.172, -.197, .551, 1.0000, .046, .044,
.086, -.129, .056, .046, 1.0000, .521,
.103, -.102, -.002, .044, .521, 1.000
  ),
  nrow = 6, ncol = 6, byrow = TRUE,
  dimnames = list(
c("JDM", "IDM", "MatWarmth", "PatWarmth", "CE_2", "CE_3"),
c("JDM", "IDM", "MatWarmth", "PatWarmth", "CE_2", "CE_3")
  )
)

cov_matrix_parenting_CE <- cor_matrix_parenting_CE * (sd_vars %*% t(sd_vars))

parenting_CE_data <- mvrnorm(n = 668, mu = mean_vars, Sigma = cov_matrix_parenting_CE)

df_civicengagement <- cbind(df_civicengagement, parenting_CE_data)

```


Great! Now I have a dataframe that includes simulated values for all of my main variables of interest. However, my simulation did not account for differences among the recruitment groups. I need to adjust my simulation to account for observed geographic/cultural differences in the parenting and civic engagement variables.  


