---
title: "Portfolio 6: Simulating Parenting and Civic Engagement Data"
---

## Overview:

#### Goal:

I want to use my portfolio to share some of the analytic work I did for my Master's thesis. However, I need to keep my data private. My goal for this piece is to simulate a dataset that resembles my thesis data. I plan to use this simulated dataset in future portfolio pieces.

#### Data:

My original dataset contained youth-reported measures of parenting practices (warmth and support for autonomy), civic engagement, and demographic information collected over three waves of a longitudinal study. Participants made up 11 different cultural/geographic groups, which were represented by effect-coded predictors in my final regression models. There was quite a bit of missing data across waves, so to simplify my simulation, I will use the most restricted subsample I had (668 participants).

#### Product:

A simulated dataset containing all variables necessary for the analyses I completed for my thesis.

#### Interpretation:

**Fill in after completing

----

### Load packages and data

```{r load-packages, message=FALSE, warning=FALSE}
library(tidyverse) 
library(ggplot2)
library(MASS)
```

### Creating the dataframe and simulating demographic information 

#### Age at Time 1 

```{r age}
df_civicengagement <- data.frame(
  id = 1:668,
  age = rnorm(668, mean = 15.61, sd = .898)
)
```

#### Geographic/cultural recruitment group 

```{r IDN}
df_civicengagement$IDN <- rep(c("Italy-Naples", "Italy-Rome", "Kenya", "Philippines", "Thailand", "Sweden", "US-AA", "US-EA", "US-H", "Colombia", "Jordan"),
  times = c(76, 84, 41, 67, 50, 33, 49, 66, 39, 74, 89),
  length.out = 668
)
```

#### Gender 

```{r gender}
df_civicengagement$male <- rep(c("not male", "male"),
  times = c(354, 314),
  length.out = 668
)
```

#### Maximum level of parental education (in years)

```{r max-parental-ed}
df_civicengagement$max_parental_ed <- rnorm(668, mean = 13.91, sd = 4.21)
```



