---
title: "Portfolio 7: Regression Models and Tables with Simulated Data"
---

## Overview:

#### Goal:

In my last portfolio piece, I simulated a dataset based on the data that I used to complete my thesis. Now, I want to replicate some of my analyses with the simulated data to show my work. My analyses consisted of a long series of regression models predicting future civic engagement from perceived parenting during adolescence. I reported the results with a series of tables that I (painstakingly) created by hand. To save myself time with this kind of reporting in the future, I also want to learn how to create APA style regression tables in R with this piece. In addition to actually running the regressions with my simulated data, I want to find the best method for producing clean tables that I can use in my future work. 

#### Data:

I simulated my data in my previous portfolio piece. The core variables are listed below. Additionally, demographic variables include age, gender, and maximum level of parental education (used as a proxy for SES).

Specific core variables include:

- Joint decision making at Time 1 (JDM; a measure of moderate support for autonomy)
- Independent decision making at Time 1 (IDM; a measure of extreme support for autonomy)
- Maternal warmth at Time 1 (MatWarmth)
- Paternal warmth at Time 1 (PatWarmth)
- Civic engagement at Time 2 (CE_2)
- Civic engagement at Time 3 (CE_3)

Recruitment groups include:

- Naples, Italy
- Rome, Italy
- Kenya
- Philippines
- Thailand
- Sweden
- US: African American
- US: European American
- US: Hispanic
- Colombia
- Jordan

#### Product:

A series of regression models presented with APA style tables created in R.

#### Interpretation:

**Fill in later

#### Sources


----

### Loading packages and data

```{r load-packages, message=FALSE, warning=FALSE}
library(tidyverse)
library(QuantPsyc)
```

```{r read-data}
civicengagement_data <- read.csv("/Users/sophieboyd/Documents/GitHub/Portfolio/docs/civicengagement_sim.csv")
```

### Data cleaning

Before I can run my regression models, I need to mean-center my predictors and dummy-code gender:

```{r mean-cent}
civicengagement_data <- civicengagement_data %>%
  mutate(age_cent = age - 15.62) %>%
  mutate(MPE_cent = max_parental_ed - 13.95) %>%
  mutate(JDM_cent = JDM - .95) %>%
  mutate(IDM_cent = IDM - .26) %>%
  mutate(MatWarmth_cent = MatWarmth - 3.46) %>%
  mutate(PatWarmth_cent = PatWarmth - 3.24) %>%
  mutate(CE_2_cent = CE_2 - 2.03) %>%
  mutate(dummy_male = case_when(
    male == "not male" ~ 0,
    male == "male" ~ 1
  ))
```

### Running the models

#### Model 1: predicting civic engagement at Time 2 from joint decision making, maternal warmth, and their interaction

```{r model1}
model1 <- lm(CE_2 ~ JDM_cent + MatWarmth_cent + MatWarmth_cent:JDM_cent , data = civicengagement_data)

summary(model1)
lm.beta(model1)
```

