---
title: "Portfolio 7: Regression Models and Tables with Simulated Data"


---

## Overview:

#### Goal:

In my last portfolio piece, I simulated a dataset based on the data that I used to complete my thesis. Now, I want to replicate some of my analyses with the simulated data to show my work. My analyses consisted of a long series of regression models predicting future civic engagement from perceived parenting during adolescence. I reported the results with a series of tables that I (painstakingly) created by hand. To save myself time with this kind of reporting in the future, I want to learn how to create APA style regression tables in R. In addition to actually running the regressions with my simulated data, my goal for this piece is to find the best method for producing clean tables that I can use in my future work. 

#### Data:

I simulated my data in my previous portfolio piece. The core variables are listed below. Additionally, demographic variables include age, gender, and maximum level of parental education (used as a proxy for SES).

Specific core variables include:

- Joint decision making at Time 1 (JDM; a measure of moderate support for autonomy)
- Independent decision making at Time 1 (IDM; a measure of extreme support for autonomy)
- Maternal warmth at Time 1 (MatWarmth)
- Paternal warmth at Time 1 (PatWarmth)
- Civic engagement at Time 2 (CE_2)
- Civic engagement at Time 3 (CE_3)

Recruitment groups include:

- Naples, Italy
- Rome, Italy
- Kenya
- Philippines
- Thailand
- Sweden
- US: African American
- US: European American
- US: Hispanic
- Colombia
- Jordan

#### Product:

A series of regression models presented with APA style tables created in R.

#### Interpretation:

In my opinion, the rempsyc package made the nicest looking tables and was the most user-friendly. This will be a very useful tool in my future work!

#### Sources
https://cran.r-project.org/web/packages/apaTables/apaTables.pdf

https://cran.r-project.org/web/packages/rempsyc/vignettes/table.html

https://github.com/crsh/papaja

----

### Loading packages and data

```{r load-packages, message=FALSE, warning=FALSE}
library(tidyverse)
library(QuantPsyc)
```

```{r read-data}
civicengagement_data <- read.csv("/Users/sophieboyd/Documents/GitHub/Portfolio/docs/civicengagement_sim.csv")
```

### Data cleaning

Before I can run my regression models, I need to mean-center my predictors and dummy-code gender:

```{r mean-cent}
civicengagement_data <- civicengagement_data %>%
  mutate(JDM_cent = JDM - .95) %>%
  mutate(IDM_cent = IDM - .26) %>%
  mutate(MatWarmth_cent = MatWarmth - 3.46) %>%
  mutate(PatWarmth_cent = PatWarmth - 3.24) %>%
  mutate(CE_2_cent = CE_2 - 2.03) %>%
  mutate(dummy_male = case_when(
    male == "not male" ~ 0,
    male == "male" ~ 1
  ))
```

### Running the models

#### Model 1: predicting civic engagement at Time 2 from joint decision making, maternal warmth, and their interaction

```{r model1}
model1 <- lm(CE_2 ~ JDM_cent + MatWarmth_cent + MatWarmth_cent:JDM_cent , data = civicengagement_data)

summary(model1)
lm.beta(model1)
```

#### Model 2: predicting civic engagement at Time 2 from joint decision making, paternal warmth, and their interaction

```{r model2}
model2 <- lm(CE_2 ~ JDM_cent + PatWarmth_cent + PatWarmth_cent:JDM_cent , data = civicengagement_data)

summary(model2)
lm.beta(model2)
```

#### Model 3: predicting civic engagement at Time 3 from joint decision making, maternal warmth, and their interaction (controlling for civic engagement at Time 2)

```{r model3}
model3 <- lm(CE_3 ~ JDM_cent + MatWarmth_cent + MatWarmth_cent:JDM_cent + CE_2_cent, data = civicengagement_data)

summary(model3)
lm.beta(model3)
```

#### Model 4: predicting civic engagement at Time 3 from joint decision making, paternal warmth, and their interaction (controlling for civic engagement at Time 2)

```{r model4}
model4 <- lm(CE_3 ~ JDM_cent + PatWarmth_cent + PatWarmth_cent:JDM_cent + CE_2_cent, data = civicengagement_data)

summary(model4)
lm.beta(model4)
```

#### Model 5: predicting civic engagement at Time 2 from independent decision making, maternal warmth, and their interaction 

```{r model5}
model5 <- lm(CE_2 ~ IDM_cent + MatWarmth_cent + MatWarmth_cent:IDM_cent, data = civicengagement_data)

summary(model5)
lm.beta(model5)
```

#### Model 6: predicting civic engagement at Time 2 from independent decision making, paternal warmth, and their interaction 

```{r model6}
model6 <- lm(CE_2 ~ IDM_cent + PatWarmth_cent + PatWarmth_cent:IDM_cent, data = civicengagement_data)

summary(model6)
lm.beta(model6)
```

#### Model 7: predicting civic engagement at Time 3 from independent decision making, maternal warmth, and their interaction (controlling for civic engagement at Time 2)

```{r model7}
model7 <- lm(CE_3 ~ IDM_cent + MatWarmth_cent + MatWarmth_cent:IDM_cent + CE_2_cent, data = civicengagement_data)

summary(model7)
lm.beta(model7)
```

#### Model 8: predicting civic engagement at Time 3 from independent decision making, paternal warmth, and their interaction (controlling for civic engagement at Time 2)

```{r model8}
model8 <- lm(CE_3 ~ IDM_cent + PatWarmth_cent + PatWarmth_cent:IDM_cent + CE_2_cent, data = civicengagement_data)

summary(model8)
lm.beta(model8)
```

### Testing out tables

Now that I've run my regressions, I am going to experiment with displaying my results in tables. I came across the following recommended packages:

- apaTables
- remPsyc
- papaja

### Loading the new packages

```{r table-packages, warning=FALSE}
library(apaTables)
library(rempsyc)
library(broom)
library(papaja)
```

### Testing each package with Model 1:

#### apaTables

```{r model1-apatables}
apa.reg.table(model1, filename = "model1table.doc")
```

I like the way that the table turned out, but saving the tables to separate word documents would be a bit inconvenient for creating multiple tables and experimenting with different layouts/aesthetic components.

#### rempsyc

```{r model1-rempsyc}
(model1table <- tidy(model1))

nice_table(model1table, broom = "lm")
```

I like the look of this one! I just need to find a way to change the labels.

#### papaja

```{r model1-papaja}
apa_model1 <- apa_print(model1)

apa_model1
```

This package requires opening a new R Markdown file with an APA template.

I am happiest with the results and functionality of the rempsyc package. Now, I need to edit the labels and create tables for each of the eight models I ran.

### Renaming the variables

I need to give the same name to two different versions of a variable, so I am renaming in two different dataframes. (Specifically, civic engagement at Time 2 is the outcome in the first four models and the centered version of the same variable is a predictor in the next four models.)

```{r rename}
civicengagement_data1 <- civicengagement_data %>%
  rename("Joint Decision Making" = JDM_cent,
         "Independent Decision Making" = IDM_cent, 
         "Maternal Warmth" = MatWarmth_cent,
         "Paternal Warmth" = PatWarmth_cent,
         "Civic Engagement at Time 2" = CE_2,
         "Civic Engagement at Time 3" = CE_3)

civicengagement_data2 <- civicengagement_data %>%
  rename("Joint Decision Making" = JDM_cent,
         "Independent Decision Making" = IDM_cent, 
         "Maternal Warmth" = MatWarmth_cent,
         "Paternal Warmth" = PatWarmth_cent,
         "Civic Engagement at Time 2" = CE_2_cent,
         "Civic Engagement at Time 3" = CE_3)
```


### Model 1

```{r model1-table}
model1 <- lm(`Civic Engagement at Time 2` ~ `Maternal Warmth` + `Joint Decision Making` + `Maternal Warmth`:`Joint Decision Making`, data = civicengagement_data1)

(model1table <- tidy(model1))

nice_table(model1table, broom = "lm")
```

I don't love that the back ticks show up on the table. I am going to rename the variables to abbreviations and add a caption to define them instead.

```{r rename2}
civicengagement_data1 <- civicengagement_data %>%
  rename("J_D_M" = JDM,
         "I_D_M" = IDM,
         "JDM" = JDM_cent,
         "IDM" = IDM_cent, 
         "MW" = MatWarmth_cent,
         "PW" = PatWarmth_cent,
         "CE2" = CE_2,
         "CE3" = CE_3)

civicengagement_data2 <- civicengagement_data %>%
  rename("J_D_M" = JDM,
         "I_D_M" = IDM,
         "JDM" = JDM_cent,
         "IDM" = IDM_cent, 
         "MW" = MatWarmth_cent,
         "PW" = PatWarmth_cent,
         "CE2" = CE_2_cent,
         "CE3" = CE_3)
```

Adding a title and notes as well:

```{r model1-table-newlabel}
model1 <- lm(CE2 ~ MW + JDM + MW:JDM, data = civicengagement_data1)

(model1table <- tidy(model1))

nice_table(model1table, broom = "lm",
           title = c("Table 1", "Predicting Civic Engagement (Time 2) from Maternal Warmth and Joint Deicision Making."),
           note = c("MW = Maternal Warmth, JDM = Joint Decision Making",
                    "* p < .05, ** p < .01, *** p < .001")
           )
```

I'm happy with this layout. Time to recreate it for the rest of the models:

### Model 2

```{r model2-table}
model2 <- lm(CE2 ~ PW + JDM + PW:JDM, data = civicengagement_data1)

(model2table <- tidy(model2))

nice_table(model2table, broom = "lm",
           title = c("Table 2", "Predicting Civic Engagement (Time 2) from Paternal Warmth and Joint Deicision Making."),
           note = c("PW = Paternal Warmth, JDM = Joint Decision Making",
                    "* p < .05, ** p < .01, *** p < .001")
           )
```

```{r model3-table}
model3 <- lm(CE3 ~ MW + JDM + MW:JDM + CE2, data = civicengagement_data2)

(model3table <- tidy(model3))

nice_table(model3table, broom = "lm",
           title = c("Table 3", "Predicting Civic Engagement (Time 3) from Maternal Warmth and Joint Deicision Making."),
           note = c("MW = Maternal Warmth, JDM = Joint Decision Making, CE2 = Civic Engagement at Time 2 (control)",
                    "* p < .05, ** p < .01, *** p < .001")
           )
```

```{r model4-table}
model4 <- lm(CE3 ~ PW + JDM + PW:JDM + CE2, data = civicengagement_data2)

(model4table <- tidy(model4))

nice_table(model4table, broom = "lm",
           title = c("Table 4", "Predicting Civic Engagement (Time 3) from Paternal Warmth and Joint Deicision Making."),
           note = c("PW = Paternal Warmth, JDM = Joint Decision Making, CE2 = Civic Engagement at Time 2 (control)",
                    "* p < .05, ** p < .01, *** p < .001")
           )
```

```{r model5-table}
model5 <- lm(CE2 ~ MW + IDM + MW:IDM, data = civicengagement_data1)

(model5table <- tidy(model5))

nice_table(model5table, broom = "lm",
           title = c("Table 5", "Predicting Civic Engagement (Time 2) from Maternal Warmth and Independent Deicision Making."),
           note = c("MW = Maternal Warmth, IDM = Independent Decision Making",
                    "* p < .05, ** p < .01, *** p < .001")
           )
```

```{r model6-table}
model6 <- lm(CE2 ~ PW + IDM + PW:IDM, data = civicengagement_data1)

(model6table <- tidy(model6))

nice_table(model6table, broom = "lm",
           title = c("Table 6", "Predicting Civic Engagement (Time 2) from Paternal Warmth and Independent Deicision Making."),
           note = c("PW = Paternal Warmth, IDM = Independent Decision Making",
                    "* p < .05, ** p < .01, *** p < .001")
           )
```

```{r model7-table}
model7 <- lm(CE3 ~ MW + IDM + MW:IDM + CE2, data = civicengagement_data2)

(model7table <- tidy(model7))

nice_table(model7table, broom = "lm",
           title = c("Table 7", "Predicting Civic Engagement (Time 3) from Maternal Warmth and Independent Deicision Making."),
           note = c("MW = Maternal Warmth, IDM = Independent Decision Making, CE2 = Civic Engagement at Time 2 (control)",
                    "* p < .05, ** p < .01, *** p < .001")
           )
```

```{r model8-table}
model8 <- lm(CE3 ~ PW + IDM + PW:IDM + CE2, data = civicengagement_data2)

(model8table <- tidy(model8))

nice_table(model8table, broom = "lm",
           title = c("Table 8", "Predicting Civic Engagement (Time 3) from Paternal Warmth and Independent Decision Making."),
           note = c("PW = Paternal Warmth, IDM = Independent Decision Making, CE2 = Civic Engagement at Time 2 (control)",
                    "* p < .05, ** p < .01, *** p < .001")
           )
```
