---
title: "Simulating and Modeling Two-Way Interactions"
---

## Overview:

#### Goal:

In my last two portfolio pieces, I focused on modeling main effects. I want to build upon this work and learn how to (1) simulate and (2) visualize data with two-way interactions. 

#### Data:

I plan to simulate a dataset with variables that could be analyzed with a 2x2 design. The dataset is inspired by my first year project, in which I presented middle school teachers with hypothetical lesson topics (all related to contemporary social issues) and asked for their thoughts/opinions about them.

The two conditions:
- Identity: whether the lesson topic focused on race or gender
- Framing: whether the lesson topic focused on social inequality or the social construction of identity

I employed a mixed design, such that identity varied between subjects and framing differed within subjects. 

Each teacher viewed four lesson topics in the actual study, but to simplify my simulation, I will give each teacher two lesson topics with the same identity condition and one of each framing condition per participant.

The main outcomes of interest (all measured on a scale from 1 to 7):
- Anticipated responses (positive/negative) from students, parents, and administrators
- Perceived appropriateness of the topics
- Perceived importance of the topics
- Perceived relevance of the topics (to the teacher's course)
- How confident the teachers felt in their ability to implement the lessons effectively
- How likely the teachers would be to introduce the topics to their students

#### Product:

A simulated dataset with my variables of interest and a series of interaction plots to visualize the results.

#### Interpretation:

** fill in after completing

#### Sources

https://www.crumplab.com/programmingforpsych/simulating-and-analyzing-data-in-r.html#simulating-data-for-multi-factor-designs 

----

### Loading packages and data

```{r load-packages, message=FALSE, warning=FALSE}
library(tidyverse) 
library(ggplot2)
library(MASS)
```

### Creating the dataframe and assigning ID numbers (sample size = 520)

```{r df}
set.seed(123)
df_lessons <- data.frame(
  id = 1:520)
```

### Creating the Identity condition (between-subjects)

```{r identity}
set.seed(123)
df_lessons$identity <- rep(c("race", "gender"),
  length.out = 520
)
```

### Creating the Framing condition (within-subjects)

This variable will be a bit more challenging to simulate. I don't know the most efficient way to achieve the data structure that I want, but in the actual project, I used a wide-to-long transformation. I am going to try to recreate the process here:

```{r framing}
set.seed(123)
df_lessons$topic1 <- rep(c("inequality"),
                           length.out = 520
)

df_lessons$topic2 <- rep(c("social construction"),
                           length.out = 520
)

df_lessons_long <- df_lessons %>%
  pivot_longer(
    cols = starts_with(c("topic")),
    names_to = "topic_number",
    values_to = "framing")

library(conflicted)
conflict_prefer("select", "dplyr")
df_lessons_long <- df_lessons_long %>%
  select(id, identity, framing)
```

That seemed to work! 

### Creating the outcome variables

Across all outcome variables, I saw the same general trend: teachers responded most negatively to lessons about the social construction of gender. Across the other three combinations of conditions, responses were neutral-to-positive at very similar levels to one another.

```{r dv-sim}
set.seed(123)
implementation <- rnorm(1040, c(4.2, 4.0, 4.3, 2.5), sd = 1.5)

df_lessons <- cbind(df_lessons_long, implementation)
```

### Checking the group means

Let's see if the mean implementation values across the four combinations of conditions match up with my expectations:

```{r group-means}
df_lessons %>%
  group_by(identity, framing) %>%
  select(implementation) %>%
  summarize(across(everything(), list(mean = mean)))

df_lessons_means <- df_lessons %>%
  group_by(identity, framing) %>%
  select(implementation) %>%
  summarize(across(everything(), list(mean = mean)))
```

Looks good! To go beyond descriptive analyses, I will run an ANOVA to see if the group means were significantly different.

```{r anova}
library(car)
reg <- lm(implementation ~ identity + framing + identity:framing, data = df_lessons)
Anova(reg, type = 3)
```

### Interpreting the interaction

```{r emtrends}
library(emmeans)

```


