---
title: "Portfolio 9: My Summer Read/Watch List"
---

## Overview:

#### Goal:

My best friends Laura and Gabby have great taste! With summer break around the corner, I want to select some good books to read and movies to watch during my time off from school. My goal for this piece is to develop a tool for choosing my next book/movie using my own preferences and data from Laura's Letterboxd account and Gabby's GoodReads account to generate recommendations.

#### Data:

I entered Laura and Gabby's reviews into a spreadsheet that includes the following variables:

- Title (the title of the movie/book)
- Rating (my friend's rating of each work from 1 to 5 stars)
- Type (movie or book)
- Genre (these are broad, somewhat imprecise categories. Options for movies are Comedy, Drama, Family, Thriller/Horror, and Romance. Options for books are Literary Fiction, Historical, Sci-Fi, Thriller, and Nonfiction/Memoir). 
- Length (in minutes for movies and in pages for books)

I recorded this information for 30 movies and 30 books.

#### Product

A simple function/algorithm that returns a recommendation for me based on my preferences for the type of work, the genre, the length, and my friends' ratings.

#### Interpretation

** fill in later

#### Sources

----

### Loading packages and data

```{r load-packages, warning=FALSE}
library(tidyverse)
library(readxl)
```

```{r load-data}
recs <- read_excel("LauraGabbyRecs.xlsx")
glimpse(recs)
```

#### Visualizing the data

Before creating my recommendation tool, I want to get a general sense of what movies/books are available in the dataset.

```{r movies-by-genre}
#ordering genres by frequency for both movies and books (after having run the bar graphs once)

recs$Genre_ordered <- fct_relevel(recs$Genre, 'Thriller/Horror', 'Drama', 'Family', 'Comedy', 'Romance', 'Thriller', 'Sci-Fi', 'Historical', 'Nonfiction/Memoir', 'Literary Fiction')

recs %>%
  filter(Type == "Movie") %>%
  ggplot(aes(x = Genre_ordered, fill = Genre_ordered)) +
  geom_bar() +
  scale_fill_viridis_d() +
  theme(legend.position = "none") +
  labs(x = 'Genre',
       y = 'Count',
       title = 'Movies by Genre')
```

```{r books-by-genre}
recs %>%
  filter(Type == "Book") %>%
  ggplot(aes(x = Genre_ordered, fill = Genre_ordered)) +
  geom_bar() +
  scale_fill_viridis_d() +
  theme(legend.position = "none") +
  labs(x = 'Genre',
       y = 'Count',
       title = 'Books by Genre')
```

```{r movie-length-distribution}
recs %>%
  filter(Type == "Movie") %>%
  ggplot(aes(x = Length, fill = "navy")) +
  geom_density() +
  scale_fill_viridis_d() +
  theme(legend.position = "none") +
  labs(x = 'Length',
       y = 'Count',
       title = 'Distribution of Movie Runtimes (in minutes)')
```

```{r book-rating-distribution}
recs %>%
  filter(Type == "Book") %>%
  ggplot(aes(x = Length, fill = "navy")) +
  geom_density() +
  scale_fill_viridis_d() +
  theme(legend.position = "none") +
  labs(x = 'Length',
       y = 'Count',
       title = 'Distribution of Book Lengths (in pages)')
```

### Building the recommendation tool


