---
title: "Portfolio 9: My Summer Read/Watch List"
---

## Overview:

#### Goal:

My best friends Laura and Gabby have great taste! With summer break around the corner, I want to select some good books to read and movies to watch during my time off from school. My goal for this piece is to develop a tool for choosing my next book/movie using my own preferences and data from Laura's Letterboxd account and Gabby's GoodReads account to generate recommendations.

#### Data:

I entered Laura and Gabby's reviews into a spreadsheet that includes the following variables:

- Title (the title of the movie/book)
- Rating (my friend's rating of each work from 1 to 5 stars)
- Type (movie or book)
- Genre (these are broad, somewhat imprecise categories. Options for movies are Comedy, Drama, Family, Thriller/Horror, and Romance. Options for books are Literary Fiction, Historical, Sci-Fi, Thriller, and Nonfiction/Memoir). 
- Length (in minutes for movies and in pages for books)

I recorded this information for 30 movies and 30 books.

#### Product

A simple set of codes that returns a recommendation for me based on my preferences for the type of work, the genre, the length, and my friends' ratings.

#### Interpretation

** fill in later

#### Sources

----

### Loading packages and data

```{r load-packages, warning=FALSE, message=FALSE}
library(tidyverse)
library(readxl)
```

```{r load-data}
recs <- read_excel("LauraGabbyRecs.xlsx")
glimpse(recs)
```

#### Visualizing the data

Before creating my recommendation tool, I want to get a general sense of what movies/books are available in the dataset.

```{r movies-by-genre}
#ordering genres by frequency for both movies and books (after having run the bar graphs once)

recs$Genre_ordered <- fct_relevel(recs$Genre, 'Thriller/Horror', 'Drama', 'Family', 'Comedy', 'Romance', 'Thriller', 'Sci-Fi', 'Historical', 'Nonfiction/Memoir', 'Literary Fiction')

recs %>%
  filter(Type == "Movie") %>%
  ggplot(aes(x = Genre_ordered, fill = Genre_ordered)) +
  geom_bar() +
  scale_fill_viridis_d() +
  theme(legend.position = "none") +
  labs(x = 'Genre',
       y = 'Count',
       title = 'Movies by Genre')
```

```{r books-by-genre}
recs %>%
  filter(Type == "Book") %>%
  ggplot(aes(x = Genre_ordered, fill = Genre_ordered)) +
  geom_bar() +
  scale_fill_viridis_d() +
  theme(legend.position = "none") +
  labs(x = 'Genre',
       y = 'Count',
       title = 'Books by Genre')
```

Lots of thriller options for both movies and books! 

```{r movie-length-distribution}
recs %>%
  filter(Type == "Movie") %>%
  ggplot(aes(x = Length, fill = "navy")) +
  geom_density() +
  scale_fill_viridis_d() +
  theme(legend.position = "none") +
  labs(x = 'Length',
       y = 'Count',
       title = 'Distribution of Movie Runtimes (in minutes)')
```

```{r book-rating-distribution}
recs %>%
  filter(Type == "Book") %>%
  ggplot(aes(x = Length, fill = "navy")) +
  geom_density() +
  scale_fill_viridis_d() +
  theme(legend.position = "none") +
  labs(x = 'Length',
       y = 'Count',
       title = 'Distribution of Book Lengths (in pages)')
```

### Building/testing the recommendation tool

```{r movie-selector}
recs_1 <- recs %>%
  filter(Type == "Movie") %>%
  filter(Genre == "Thriller/Horror") %>%
  filter(Rating >= 4) %>%
  filter(Length <= 120) 

sample(recs_1$Title, 1)
  
```

```{r book-selector}
recs_2 <- recs %>%
  filter(Type == "Book") %>%
  filter(Genre == "Sci-Fi") %>%
  filter(Rating >= 4) %>%
  filter(Length <= 400) 

sample(recs_2$Title, 1)
  
```

This is my first attempt at getting a random movie recommendation and book recommendation that fit my selected criteria. It is a bit inconvenient to have to save a new dataframe that meets each set of criteria, but I think it works well enough for what I need to do. I will create a more general template that I can refer back to later: 

### Step-by-step recommendation tool:

```{r tool-template, eval = FALSE}
new_df <- recs %>%
  # Movie or book?
  filter(Type == "____") %>%
  # What genre are you in the mood for?
  filter(Genre == "____") %>%
  # What is the minimum rating you trust?
  filter(Rating >= __) %>%
  # How much time do you have? (enter max length preference in minutes/pages)
  filter(Length <= ____) 

sample(new_df$Title, 1)
```

Emerson let me test out my code with her preferences too: 

```{r Emerson-pref}
EE_df <- recs %>%
  filter(Type == "Movie") %>%
  filter(Genre == "Comedy") %>%
  filter(Rating >= 3) %>%
  filter(Length <= 120)

sample(EE_df$Title, 1)
```

Her recommended movie was Y2K!

### Saving my progress

As I watch/read the recommended movies/books, I want to be able to keep track of what I have watched/read already so I can continuing narrowing my recommendations down to works that are new to me. I also want to save my own ratings of the movies and books to compare to my friends' ratings. I will create new variables in my dataframe to save my progress and filter my future recommendations:

```{r my-progress}
recs <- recs %>%
  mutate(watched = 0) %>%
  mutate(my_rating = 0)
```

### Putting it all together:

To test everything step-by-step, I am going to request a movie recommendation, watch the movie this evening, and return to update my dataset with my own rating. 

```{r my-test}
test_df <- recs %>%
  # Movie or book?
  filter(Type == "Movie") %>%
  # What genre are you in the mood for?
  filter(Genre == "Comedy") %>%
  # What is the minimum rating you trust?
  filter(Rating >= 3.5) %>%
  # How much time do you have? (enter max length preference in minutes/pages)
  filter(Length <= 120) 

sample(test_df$Title, 1)
```

Going to watch "This Is Spinal Tap" and report back!

```{r update}
  
```

