---
title: "Portfolio 10: Summer Workflow"
---

## Overview:

#### Goal:

In my last portfolio piece, I focused on what I would do with my free time this summer. For this piece, I will turn my attention to my outstanding projects that I hope to complete or push forward by the end of the summer. My goal is to create a graphic organizer to help me break the projects down into smaller tasks and keep track of my progress. 

#### Data:

No dataset for this one- I will just be entering a list of my projects/tasks.

#### Product:

** fill in after completing

#### Interpretation:

** fill in after completing

#### Sources:

https://sites.temple.edu/psmgis/2017/07/30/flow-charts-in-r-using-diagrammer/

https://rich-iannone.github.io/DiagrammeR/#:~:text=Any%20time%20we%20add%20a,edges%20and%20bind%20attribute%20data.

---- 

I am going to start by searching online for recommended packages/tools for creating graphic organizers and/or flow charts to test out. 

Options include:

- DiagrammeR
- 


```{r load-packages, warning=FALSE}
library(DiagrammeR)
```

```{r attempt1}
DiagrammeR("graph TB;
           Analysis --> Writing;
           Analysis --> Cleaning;
           Analysis --> Hypothesis_Testing;
           Writing --> Method;
           Writing --> Results;
           Writing --> Discussion")
```

