---
title: "Portfolio 10: Summer Workflow"
---

## Overview:

#### Goal:

In my last portfolio piece, I focused on what I would do with my free time this summer. For this piece, I will turn my attention to my outstanding projects that I hope to complete or push forward by the end of the summer. My goal is to create a graphic organizer to help me break the projects down into smaller tasks and keep track of my progress. 

#### Data:

No dataset for this one- I will just be entering a list of my projects/tasks.

#### Product:

** fill in after completing

#### Interpretation:

** fill in after completing

#### Sources:

https://sites.temple.edu/psmgis/2017/07/30/flow-charts-in-r-using-diagrammer/

https://rich-iannone.github.io/DiagrammeR/#:~:text=Any%20time%20we%20add%20a,edges%20and%20bind%20attribute%20data.

---- 

I am going to start by searching online for recommended packages/tools for creating graphic organizers and/or flow charts to test out. 

Options include:

- DiagrammeR
- 


```{r load-packages, warning=FALSE}
library(DiagrammeR)
```

Some starter code to see how it works:

```{r attempt1}
DiagrammeR("graph LR;
           A --> B;
           B --> C;
           B --> D")

DiagrammeR("graph TB;
           A --> B;
           B --> C;
           B --> D")
```

Based on how the levels work/look, I think it would make sense to make a separate organizer for each project I am working on. I will start with the remaining tasks for my FYP:

```{r FYP1}
DiagrammeR("graph TB;
           FYP --> Data;
           FYP --> Writing;
           Data --> Cleaning;
           Data --> Analysis;
           Writing --> Results;
           Writing --> Discussion")
```

```{r nutrition1}
DiagrammeR("graph TB;
           NP --> Writing;
           Writing --> Results3;
           Writing --> Results4")
```

```{r PAC1}
DiagrammeR("graph TB;
           PAC --> Data;
           PAC --> Writing;
           Data --> Cleaning;
           Data --> Analysis;
           Writing --> Introduction;
           Writing --> Discussion")
```

```{r thesis1}
DiagrammeR("graph TB;
           Thesis --> Planning;
           Thesis --> Data;
           Thesis --> Writing;
           Planning --> Meeting;
           Planning --> Change_groups;
           Data --> Cleaning;
           Data --> Analysis;
           Writing --> Edit_intro;
           Writing --> Edit_method;
           Writing --> Results;
           Writing --> Discussion")
```

Now that I have the basic layout of tasks that need to be completed, I want to add some sort of timeline element to the diagrams. I am going to try to color code the boxes by month to reflect when I would like each task completed:

```{r FYP2}
grViz("
  digraph {
    node [shape = circle, style = filled]
    
    A [fillcolor = Cyan]
    B [fillcolor = Gold]
    
    A -> B [color = grey]
  }
")


grViz("digraph {
node [shape = rectangle, style = filled]

FYP [fillcolor = Blue]
Data [fillcolor = Purple]
Writing [fillcolor = Green]
Cleaning [fillcolor = Purple]
Analysis [fillcolor = Purple]
Results [fillcolor = Green]
Discussion [fillcolor = Green]

           FYP -> Data [color = grey];
           FYP -> Writing [color = grey];
           Data -> Cleaning [color = grey];
           Data -> Analysis [color = grey];
           Writing -> Results [color = grey];
           Writing -> Discussion [color = grey]

}")
          
```

